---
---
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)
library(dplyr)
library(ggplot2)
```

## Datenaufbereitung

```{r Datenaufbereitung, include=FALSE}
# Daten einlesen
df <- read_dta("Full_dset.dta")

# Vereinheitlichung und Filterung der Daten
df_clean <- df %>% 
  rename(
    vote_for = V202073,  # Abstimmung: 1 = Biden, 2 = Trump
    age = V201507x,      # Alter
    sex = V201600,       # Geschlecht: 1 = Männlich, 2 = Weiblich
    education = V201510, # Bildungsniveau
    income = V201617x,    # Einkommen
    race = V201549x
  ) %>% 
  filter(vote_for == 1 | vote_for == 2) %>%
  filter(sex == 1 | sex == 2)

# Variablen transformieren
df_clean <- df_clean %>% 
  mutate(
    vote_trump1 = ifelse(vote_for == 2, 1, 0),  # 1 = Trump, 0 = Biden
    sex_male = ifelse(sex == 1, 1, 0)
  )


```

## Hypothese 1

```{r H1_glm_model}
h1 <- glm(vote_trump1 ~ sex_male, data = df_clean, family = binomial)
summary(h1)

```

```{r, warning = FALSE}
ggplot(df_clean, aes(sex_male, vote_trump1)) + 
  geom_point() +
  geom_smooth(method = "glm", col = "red") +
  labs(
    x = "Geschlecht (1 = Männlich, 0 = Weiblich)", 
    y = "Trump-Wahl (1 = Ja, 0 = Nein)",  
    title = "Zusammenhang zwischen Geschlecht und Trump-Wahl"
  )

```

Checking for confounder race -\> doesn't change coefficient for sex_male

Barplot

```{r, warning = FALSE}
# Data preparation with German labels
bar_data <- df_clean %>%
  group_by(sex_male) %>%
  summarise(percentage = mean(vote_trump1) * 100) %>%
  mutate(sex = ifelse(sex_male == 1, "Männlich", "Weiblich"))  # German labels

# Create the bar plot with updated styles
ggplot(bar_data, aes(x = sex, y = percentage, fill = sex)) +
  geom_bar(stat = "identity", color = "white", width = 0.4) +  # Bar plot with narrower bars
  scale_fill_manual(values = c("Männlich" = "#224598", "Weiblich" = "#B22032")) +  # Custom colors
  labs(
    title = "Anteil von Männern und Frauen, die für Trump stimmten",
    x = "Geschlecht",
    y = "Anteil (%)"
  ) +
  scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +  # Y-axis scale 0 to 100
  theme_minimal(base_family = "Arial") +  # Minimal theme with base font
  theme(
    plot.background = element_rect(fill = "#D9D9D9", color = NA),  # Background color
    panel.grid.major = element_line(color = "#A0A0A0", size = 0.5),  # Add major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    panel.background = element_rect(fill = "#D9D9D9", color = NA),  # Panel background
    axis.title = element_text(size = 14, color = "#040041"),  # Axis titles in dark blue
    axis.text = element_text(size = 12, color = "#040041"),   # Axis text (numbers) in dark blue
    axis.ticks = element_line(color = "#040041"),  # Axis ticks in dark blue
    plot.title = element_text(size = 16, hjust = 0.5, color = "#040041"),  # Title in dark blue
    legend.position = "none"  # Remove legend
  )



```

```{r}
h1_checkrace <- glm(vote_trump1 ~ sex_male + race, data = df_clean, family = binomial)
summary(h1_checkrace)

```

Calculate probability -\> probability to vote for Trump 28% higher for male than female

```{r}
log_odds_female <- predict(h1, newdata = data.frame(sex_male = 0), type = "link")

# Predicted log-odds for males (sex_male = 1)
log_odds_male <- predict(h1, newdata = data.frame(sex_male = 1), type = "link")

# Convert log-odds to odds
odds_female <- exp(log_odds_female)
odds_male <- exp(log_odds_male)

# Calculate odds ratio
odds_ratio <- odds_male / odds_female
print(odds_ratio)
```

## Hypothese 2

Datenaufbereitung

```{r}
# Einfluss des Bildungsniveaus auf die Trump-Wahl, moderiert durch Geschlecht
# Filter out invalid categories (keep only values between 0 and 10)
df_clean_edu <- df_clean[df_clean$education >= 0 & df_clean$education <= 10, ]

# Create a new categorical variable by grouping the education levels
df_clean_edu$education_category <- cut(
  df_clean_edu$education,
  breaks = c(0, 3, 5, 8),  # Define category boundaries
  labels = c("1-3", "4-5", "6-8"),  # Define category labels
  right = TRUE  # Include upper bound in each interval
)
```

Check education distribution

```{r}
table(df_clean_edu$education_category)
```

H2 glm model

```{r}
h2 <- glm(vote_trump1 ~ education * sex_male, data = df_clean_edu)
summary(h2)
```

Calculate probability for male

```{r}
# Create a data frame for different combinations of sex_male and education
new_data <- data.frame(
  sex_male = 1,                         # Males
  education = c(1, 2, 3)                # Education levels: 1 (High school), 2 (College), 3 (University)
)

# Predict log-odds using the model
log_odds <- predict(h2, newdata = new_data, type = "link")

# Convert log-odds to probabilities
probabilities <- 1 / (1 + exp(-log_odds))

# Combine results for better understanding
results <- cbind(new_data, log_odds, probabilities)
print(results)
```

Calculate probability for female -\> probability dicreases faster for females, effect of high education stronger for male

```{r}
# Create a data frame for different combinations of sex_male and education
new_data <- data.frame(
  sex_male = 0,                         # Males
  education = c(1, 2, 3)                # Education levels: 1 (High school), 2 (College), 3 (University)
)

# Predict log-odds using the model
log_odds <- predict(h2, newdata = new_data, type = "link")

# Convert log-odds to probabilities
probabilities <- 1 / (1 + exp(-log_odds))

# Combine results for better understanding
results <- cbind(new_data, log_odds, probabilities)
print(results)
```

Graph

```{r, warning = FALSE}
ggplot(df_clean_edu, aes(x = education, y = vote_trump1, color = factor(sex_male))) + 
  geom_smooth(
    method = "glm", 
    method.args = list(family = binomial), 
    se = FALSE, 
    size = 1.2  # Thicker line for better visibility
  ) + 
  scale_color_manual(
    values = c("1" = "#224598", "0" = "#B22032"),  # Dark blue for male, red for female
    labels = c("Mann", "Frau")
  ) +
  labs(
    title = "Bildungsniveau und Trump-Wahl nach Geschlecht",
    x = "Bildungsniveau", 
    y = "Trump-Wahl (1 = Ja, 0 = Nein)",
    color = "Geschlecht"
  ) +
  theme_minimal(base_family = "Arial") +  # Minimal theme with consistent font
  theme(
    plot.background = element_rect(fill = "#D9D9D9", color = NA),  # Light gray background
    panel.background = element_rect(fill = "#D9D9D9", color = NA),  # Panel background
    panel.grid = element_line(color = "#B0B0B0"),  # Light gray grid lines
    axis.title = element_text(size = 14, color = "#040041"),  # Axis titles in dark blue
    axis.text = element_text(size = 12, color = "#040041"),   # Axis text (numbers) in dark blue
    axis.ticks = element_line(color = "#040041"),  # Axis ticks in dark blue
    plot.title = element_text(size = 16, hjust = 0.5, color = "#040041"),  # Title in dark blue
    legend.position = "top",  # Position legend at the top
    legend.text = element_text(color = "#040041", size = 12),  # Legend text color and size
    legend.title = element_text(color = "#040041", size = 12)  # Legend title in dark blue
  )

```

## Hypothese 3

Datenaufbereitung

```{r}
df_edu_inc <- df_clean %>% 
  filter(education >= 1 & education <= 8) %>% 
  filter(income >= 1 & income <= 22) %>%
  mutate(income2 = income^2)
```

Model 3

```{r}
h3_income <- glm(vote_trump1 ~ income + I(income^2), data = df_edu_inc, family = binomial)
summary(h3_income)
```

Graph 1

```{r, warning = FALSE}
ggplot(df_edu_inc, aes(x = income, y = vote_trump1)) +
  geom_smooth(
    method = "glm", 
    method.args = list(family = binomial), 
    formula = y ~ x + I(x^2),  # Hier wird das quadratische Einkommen hinzugefügt
    se = FALSE, 
    color = "blue"
  ) +
  labs(
    title = "Einfluss von Einkommen auf die Trump-Wahl",
    x = "Einkommen",
    y = "Wahrscheinlichkeit für Trump-Wahl"
  ) +
  theme_minimal()
```

Graph 2

```{r, warning = FALSE}
# Einkommensgruppen sortieren: Mittleres Einkommen in der Mitte
df_edu_inc <- df_edu_inc %>% 
  mutate(
    income_group = factor(case_when(
      income <= 7 ~ "Niedrig",
      income <= 14 ~ "Mittel",
      TRUE ~ "Hoch"
    ), levels = c("Niedrig", "Mittel", "Hoch")) # Reihenfolge festlegen
  )

# Graph erstellen
ggplot(df_edu_inc, aes(x = income_group, y = vote_trump1, fill = income_group)) +
  stat_summary(
    fun = mean, geom = "bar", color = "white", width = 0.7
  ) +
  scale_fill_manual(
    values = c("Hoch" = "#B22032", "Mittel" = "#5F9EA0", "Niedrig" = "#224598")
  ) +
  labs(
    title = "Trump-Wahl nach Einkommensgruppen",
    x = "Einkommensgruppe",
    y = "Wahrscheinlichkeit Trump-Wahl"
  ) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#D9D9D9", color = NA),
    panel.background = element_rect(fill = "#D9D9D9", color = NA),
    axis.text = element_text(size = 12, color = "#040041"),
    axis.title = element_text(size = 14, color = "#040041"),
    plot.title = element_text(size = 16, hjust = 0.5, color = "#040041"), # "face" entfernt
    plot.subtitle = element_text(size = 14, hjust = 0.5, color = "#040041"),
    panel.grid = element_line(color = "#B0B0B0"),
    legend.position = "none" # Legende entfernen
  )


```

Check race

```{r}
h3_income_checkrace <- glm(vote_trump1 ~ income + I(income^2) + race, data = df_edu_inc, family = binomial)
summary(h3_income_checkrace)
```
